---
title: "CEO Compensation regression analysis"
author: "Afsar Ali"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
---

## Code header 

```{r}
# Course: BUAN 5210
# Title: Final Project EDA
# Purpose: CEO Compensation regression analysis
# Date: Dec 1st, 2017
# Author: Afsar Ali
```


```{r message=FALSE, warning=FALSE}
# Clear working environment
rm(list=ls(all=TRUE)) 

# The tidyverse package contains ggplot2, tibble, tidyr, readr, purr, and dplyr among others
library(tidyverse)
# The gridExtra package contains grid.arrange function used to combine plots
library(gridExtra)
# The GGally package contains ggpairs which is a custom correlation graph for ggplot2
library(GGally)
# Load gridExtra so can plot more than one graph with grid.arrange
library(gridExtra)
#load other packages
library(plotly)
library(psych)
library(knitr)
library(htmlTable)

```


```{r message=FALSE, warning=FALSE}
#Load and Review data
dat <- read.csv("ECON5100_project_data.csv", header = TRUE)
#datd <- read.csv("ECON5100_project_data_definition.csv", header = TRUE) 
dat <- mutate(dat, DRatio = (DLTT/AT)) #Calculating long term debt ratio

#Make a subset of relevent data
dat1 <- select(dat, "AT","NI", "CH", "DLTT", "PRCH_C", "PRCL_C", "DRatio", "TDC2", "SALARY", "BONUS", "OTHCOMP", "AGE", "GENDER")
dat1<-  mutate (dat1, ageg = cut(AGE, 6)) #grouping age by 6
datm <- dat1 %>%  filter(GENDER =="MALE") #nale only 
datf <- dat1 %>%  filter(GENDER =="FEMALE") #female only

#attach the file for use
attach(dat1)

glimpse(dat)
summary(dat)
```

## Histogram of AGE

Mostly middle aged  

```{r message=FALSE, warning=FALSE}
plot_ly(x = ageg,
             type = "histogram")

plot_ly(x = ageg,
             type = "histogram",
             histnorm = "probability")
```

## Histrogram of Gender

lot of Males (93% is Male)
Mostly Middle aged Men

```{r message=FALSE, warning=FALSE}
plot_ly(x = GENDER,
             type = "histogram")

plot_ly(x = GENDER,
             type = "histogram",
             histnorm = "probability")
```

## Histogram of other Variables 

Nothing Significant From these Histogram

```{r message=FALSE, warning=FALSE}
plot_ly(x = CH,
             type = "histogram") #Total Cash
plot_ly(x = AT,
             type = "histogram") #Assets - Total
plot_ly(x = NI,
             type = "histogram") #Net Income (Loss)
plot_ly(x = DLTT,
             type = "histogram") #Long-Term Debt - Total
plot_ly(x = PRCH_C,
             type = "histogram") #Price High - Annual - Calendar
plot_ly(x = PRCL_C,
             type = "histogram") #Price Low - Annual - Calendar
plot_ly(x = DRatio,
             type = "histogram") #Debt Ratio
plot_ly(x = SALARY,
             type = "histogram") #Salary
plot_ly(x = BONUS,
             type = "histogram") #The dollar value of a bonus earned by the named executive officer during the fiscal year.
plot_ly(x = OTHCOMP,
             type = "histogram") #All Other Compensation ($)
plot_ly(x = TDC2,
             type = "histogram") #Total Compensation (Salary + Bonus + Other Annual + Restricted Stock Grants + LTIP Payouts + All Other + Value of Options Exercised).

```

## Histogram of other Variables using log to find significance 

Normal Distribution

```{r message=FALSE, warning=FALSE}
plot_ly(x = log(CH),
             type = "histogram")
plot_ly(x = log(AT),
             type = "histogram")
plot_ly(x = log(NI),
             type = "histogram")
plot_ly(x = log(DLTT),
             type = "histogram")
plot_ly(x = log(PRCH_C),
             type = "histogram")
plot_ly(x = log(PRCL_C),
             type = "histogram")
plot_ly(x = log(SALARY),
             type = "histogram")
plot_ly(x = log(BONUS),
             type = "histogram")
plot_ly(x = log(OTHCOMP),
             type = "histogram")
plot_ly(x = log(TDC2),
             type = "histogram")
plot_ly(x = log(DRatio),
             type = "histogram")
```

##Relationships between selected variables

High to Low relationships
-AT and DLTT, SALARY, BONUS
-NI and TDC2, SALARY, BONUS
-CH and SALARY, BONUS, TDC2
-DLTT and SALARY
-PRCH_C, PRCH_L and TDC2
-SALARY and AGE



```{r message=FALSE, warning=FALSE}
par(mfrow=c(1,1))
pairs.panels(dat1)

```

-No Significant Difference observed on Male only Data probably due to high number of observation 

```{r message=FALSE, warning=FALSE}
#observe male data only 
pairs.panels(datm)

```

-AT and Bonus has a huge difference between Male .21 and female .60
-DLTT and Bonus has a huge difference between Male .05 and female .56
-TDC2 and CH has a small difference between Male .25 and female .36
-NI and Stock price has difference between Male and female
-All other relationships are high compared to Men Probably due to very low observations

Does female Execs has to perform better then men in average? 
Does firm have better with female Exec?

```{r message=FALSE, warning=FALSE}
#observe only  female data
pairs.panels(datf)
```

```{r message=FALSE, warning=FALSE}
#Getting Ideas about graphs
dat1 %>%
  select("AT", "NI", "CH","TDC2", "SALARY", "BONUS", "AGE", "GENDER") %>%
ggpairs(dat1)
```



## Compbining Datset

Added age group to my data

```{rmessage=FALSE, warning=FALSE}

hist(dat$AGE)
```

```{rmessage=FALSE, warning=FALSE}
group_by(dat1, AGE)
summarize(group_by(dat1, AGE), aSalary=mean(SALARY))
```


## Probability Stats

Added age group to my data
Review Probability


```{r message=FALSE, warning=FALSE}
#Review Probability
htmlTable(as.data.frame(round(prop.table(xtabs(~ GENDER + ageg,  
      data = dat1 %>%
      mutate(AGE = AGE) %>% 
      mutate(ageg = cut(AGE, 6))), 2), 2)),
      caption = "Frequency of Price Difference",
      header = paste(c("Brand", "Price Difference", "&dagger; Frequency")),
      tfoot="&dagger; Data from Midterm Project")
##added new variable ageg (AGE GROUP Based on this)

```


Female low probability

```{r message=FALSE, warning=FALSE}
(d <- as.data.frame(prop.table(xtabs(~ GENDER + ageg,  
  data = dat1 %>%
  mutate (ageg = cut(AGE, 6)) %>% # step 2
  mutate (ageg = recode(ageg, "(0,30)" = "30 and under",
                             "(31,45)" = "31-45",
                             "(46,65)" = "46-65",
                             "(66,100)" = "65+"))), 2)) %>%
# 3 and 4. Visualize conditional probability of brand by price difference type
  ggplot(aes(x = GENDER, y = Freq, fill = ageg)) +
  geom_bar(stat="identity", position = "dodge", color = "black") +
  geom_text(aes(label = round(Freq,2)), vjust=1.5, color = "white",
            position = position_dodge(0.9), size = 4)+
  ggtitle("Women are less likely to be Executives", sub = "Probabbility Grouped by Age")
)

```

## EDA Graphs

Middle age Men has highest number Bonus and Salary


```{r message=FALSE, warning=FALSE}
ggplot(dat, aes(x = AGE, y = BONUS, color = GENDER)) +
  geom_point() +
  ggtitle("Middle age men has highest number of bonus with outliers", sub = "Small number of Female observations")

```

Over-all Observation

```{r message=FALSE, warning=FALSE}
grid.arrange(
ggplot(dat, aes(x = AGE, y = SALARY, color = GENDER)) +
  geom_point() +
  stat_density2d() +
  ggtitle("Mostly Middle aged Men", sub = "Small number of Female observations"),
ggplot(dat, aes(x = AGE, y = SALARY, color = GENDER)) +
  geom_point() +
  stat_density2d(aes(fill =..density..), geom = "raster", contour = FALSE) +
  ggtitle("Mostly Middle aged Men", sub = "Heat map, light color indicates more observations"),
ncol = 2
)
```


```{r message=FALSE, warning=FALSE}
grid.arrange(
dat %>%
  group_by(GENDER) %>%
  summarise(age_mean = mean(AGE)) %>%
  ggplot(aes(x = GENDER, y = age_mean, fill = GENDER)) +
    geom_bar(stat="identity") +
    coord_flip(), 
dat %>%
  ggplot(aes(x = GENDER, y = AGE, fill = GENDER)) +
    geom_boxplot() +
    coord_flip() +
    guides(fill = FALSE, ylab = FALSE),
ncol = 2
)

```

##Female Compensation in average 
=========================================
As female age they get higher compensation


```{r message=FALSE, warning=FALSE}
datf %>%
  group_by(ageg) %>%
  summarise(Avg_Com = mean(TDC2)) %>%
  ggplot(aes(x = ageg, y = Avg_Com, )) +
  geom_bar(stat="identity", position = "dodge", color = "black") +
  labs(fill = "Gender") +
# Add and improve titles and labels
    ggtitle("In Average as female age there Compensation gets Larger") + 
    ylab("Average Compensation by Thousands") +
    xlab("Age Group") +
    scale_x_discrete(labels=c("Under 40", "40 to 51", "51 to 62", "62 to 73", "73 to 84", "84 +", "NA")) +
 theme_classic() + 
  theme(axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),)
```




```{r message=FALSE, warning=FALSE}
datm %>%
  group_by(ageg) %>%
  summarise(Avg_Com = mean(TDC2)) %>%
  ggplot(aes(x = ageg, y = Avg_Com, )) +
  geom_bar(stat="identity", position = "dodge", color = "black") +
  labs(fill = "Gender") +
# Add and improve titles and labels
    ggtitle("In Average as female age there Compensation gets Larger") + 
    ylab("Average Compensation by Thousands") +
    xlab("Age Group") +
    scale_x_discrete(labels=c("Under 40", "40 to 51", "51 to 62", "62 to 73", "73 to 84", "84 +", "NA")) +
 theme_classic() + 
  theme(axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),)
  
```


```{r message=FALSE, warning=FALSE}
dat1 %>%
  group_by( GENDER ) %>%
  summarise( count = n(),
              proportion = sum(count) / nrow(dat1),
              percent = proportion * 100)

dat1 %>%
  group_by( AGE ) %>%
  ggplot() +
    geom_bar(aes(x = GENDER)) +
    ggtitle("Middle Age")

```

##Avearage Compensation by Gender
=========================================
Lets compare Total Compensation side by side   

+ We can see that: 
    - There are no Female Executive over 84
    - Female age group has a big compensation jump from 62-73 to 73-84 

```{r message=FALSE, warning=FALSE}
# Average profit by region and category 
dat1 %>%
  group_by(ageg, GENDER) %>%
  summarise(Avg_Com = mean(TDC2)) %>%
  ggplot(aes(x = ageg, y = Avg_Com, fill = reorder(GENDER, desc(Avg_Com)))) +
  geom_bar(stat="identity", position = "dodge", color = "black") +
  geom_hline(yintercept = 0, color = "black") +
  labs(fill = "Gender") +
# Add and improve titles and labels
    ggtitle("Middle age female age groups has a big compensation jump") + 
    ylab("Average Compensation by Thousands") +
    xlab("Age Group") +
    scale_x_discrete(labels=c("Under 40", "40 to 51", "51 to 62", "62 to 73", "73 to 84", "84 +", "NA")) +
 theme_classic() + 
  theme(axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),)
  
```

##Multiple Regression Analysis
Testing Different Models for Significance Analysis

1. Most statistically significant  
2. Have the largest impact  

Visualize which variables have coefficients statistically different from zero  


```{r message=FALSE, warning=FALSE}
attach(dat1)
# ANOVA
baseMod <- lm(TDC2 ~ AGE + BONUS + AT + CH + DLTT + NI + PRCH_C + PRCL_C)
mod1 <- lm(TDC2 ~ AGE + BONUS + AT + CH + DLTT + NI)
mod2 <- lm(TDC2 ~ AGE + BONUS + AT + CH + DLTT)
mod3 <- lm(TDC2 ~ AGE + BONUS + AT + CH)
mod4 <- lm(TDC2 ~ AGE + BONUS + AT)
anova(baseMod, mod1, mod2, mod3, mod4)
summary(baseMod)
```

```{r message=FALSE, warning=FALSE}
attach(datm)
# ANOVA
baseMod <- lm(TDC2 ~ AGE + BONUS + AT + CH + DLTT + NI + PRCH_C + PRCL_C)
mod1 <- lm(TDC2 ~ AGE + BONUS + AT + CH + DLTT + NI)
mod2 <- lm(TDC2 ~ AGE + BONUS + AT + CH + DLTT)
mod3 <- lm(TDC2 ~ AGE + BONUS + AT + CH)
mod4 <- lm(TDC2 ~ AGE + BONUS + AT)
anova(baseMod, mod1, mod2, mod3, mod4)
summary(baseMod)
```

```{r message=FALSE, warning=FALSE}
attach(datf)
# ANOVA
baseMod <- lm(TDC2 ~ AGE + BONUS + AT + CH + DLTT + NI + PRCH_C + PRCL_C)
mod1 <- lm(TDC2 ~ AGE + BONUS + AT + CH + DLTT + NI)
mod2 <- lm(TDC2 ~ AGE + BONUS + AT + CH + DLTT)
mod3 <- lm(TDC2 ~ AGE + BONUS + AT + CH)
mod4 <- lm(TDC2 ~ AGE + BONUS + AT)
anova(baseMod, mod1, mod2, mod3, mod4)
summary(baseMod)
```

##Visualization of Data Regression 
=========================================

Age Groups Has most Confident Interval Variance  


```{r message=FALSE, warning=FALSE}
attach(dat1)
mod <- lm(TDC2 ~ AGE + BONUS + AT + CH + DLTT + NI + PRCH_C + PRCL_C+ OTHCOMP+ ageg)
coe <- summary(mod)$coefficients # get coefficients and related stats
coe_CI <- as.data.frame(cbind(coe[-1, ], confint(mod)[-1, ])) # find and bind CI, remove Int 

# Rename results data frame
names(coe_CI) <- c("estimate", "se", "t", "pval","low_CI","high_CI") 

htmlTable(round(coe_CI[order(coe_CI$pval, decreasing=FALSE),],3))

(g1 <- ggplot(coe_CI, aes(x = estimate, y = reorder(row.names(coe_CI),desc(pval)))) +
  geom_point(size = 3) +
  xlim(min(coe_CI$low_CI), max(coe_CI$high_CI)) +
  ylab("Variable") +
  xlab("Coefficient") +
  theme_bw() +
  ggtitle("ALL Data: Age group has different confident intervals") +
  geom_segment(aes(yend = reorder(row.names(coe_CI),desc(pval))), 
               xend = coe_CI$high_CI) +
  geom_segment(aes(yend = reorder(row.names(coe_CI),desc(coe_CI$pval))), 
               xend = coe_CI$low_CI) +
   xlab("Coefficient with Confidence Interval") +
   geom_vline(xintercept = 0, color = "red")
)

```

MALE ONLY

```{r message=FALSE, warning=FALSE}
attach(datm) #Only Male Data
mod <- lm(TDC2 ~ AGE + BONUS + AT + CH + DLTT + NI + PRCH_C + PRCL_C+ OTHCOMP+ ageg)
coe <- summary(mod)$coefficients # get coefficients and related stats
coe_CI <- as.data.frame(cbind(coe[-1, ], confint(mod)[-1, ])) # find and bind CI, remove Int 

# Rename results data frame
names(coe_CI) <- c("estimate", "se", "t", "pval","low_CI","high_CI") 

htmlTable(round(coe_CI[order(coe_CI$pval, decreasing=FALSE),],3))

(g1 <- ggplot(coe_CI, aes(x = estimate, y = reorder(row.names(coe_CI),desc(pval)))) +
  geom_point(size = 3) +
  xlim(min(coe_CI$low_CI), max(coe_CI$high_CI)) +
  ylab("Variable") +
  xlab("Coefficient") +
  theme_bw() +
  ggtitle("Male only: Age group has different confident intervals") +
  geom_segment(aes(yend = reorder(row.names(coe_CI),desc(pval))), 
               xend = coe_CI$high_CI) +
  geom_segment(aes(yend = reorder(row.names(coe_CI),desc(coe_CI$pval))), 
               xend = coe_CI$low_CI) +
   xlab("Coefficient with Confidence Interval") +
   geom_vline(xintercept = 0, color = "red")
)

```

FEMALE ONLY: USING LOG

```{r message=FALSE, warning=FALSE}
attach(datf) #Only Female Data
mod <- lm(log(TDC2) ~ AGE + BONUS + AT + CH + DLTT + NI + PRCH_C + PRCL_C+ OTHCOMP+ ageg)
coe <- summary(mod)$coefficients # get coefficients and related stats
coe_CI <- as.data.frame(cbind(coe[-1, ], confint(mod)[-1, ])) # find and bind CI, remove Int 

# Rename results data frame
names(coe_CI) <- c("estimate", "se", "t", "pval","low_CI","high_CI") 

htmlTable(round(coe_CI[order(coe_CI$pval, decreasing=FALSE),],3))

(g1 <- ggplot(coe_CI, aes(x = estimate, y = reorder(row.names(coe_CI),desc(pval)))) +
  geom_point(size = 3) +
  xlim(min(coe_CI$low_CI), max(coe_CI$high_CI)) +
  ylab("Variable") +
  xlab("Coefficient") +
  theme_bw() +
  ggtitle("Female only: Age group has different confident intervals") +
  geom_segment(aes(yend = reorder(row.names(coe_CI),desc(pval))), 
               xend = coe_CI$high_CI) +
  geom_segment(aes(yend = reorder(row.names(coe_CI),desc(coe_CI$pval))), 
               xend = coe_CI$low_CI) +
   xlab("Coefficient with Confidence Interval") +
   geom_vline(xintercept = 0, color = "red")
)

```


##Chow test

```{r message=FALSE, warning=FALSE}
#== Chow test
chow.test <- local({
  m <- 2    # number of restrictions (delta_0 = delta_1 = 0)
  n <- dim(dat1)[1]
  full.1 <- lm( TDC2 ~ AGE, data=subset(dat, GENDER == "FEMALE"))  # female regression
  full.2 <- lm( TDC2 ~ AGE, data=subset(dat, GENDER == "MALE"))    # male regression
  full <- lm(TDC2 ~ AGE + GENDER + GENDER*AGE, data=dat)          # full model using dumies
  rest <- lm(TDC2 ~ AGE + AGE, data=dat)                          # restricted model
  SSE.1 <- sum( resid(full.1)^2 )
  SSE.2 <- sum( resid(full.2)^2 )
  SSE.full <- sum( resid(full)^2 )
  all.equal(SSE.full, SSE.1 + SSE.2)
  SSE.rm <- sum( resid(rest)^2 )
  chow <- (( SSE.rm - (SSE.1 + SSE.2)) / m ) / ( (SSE.1 + SSE.2) / (n - 2*m) )
  pval.chow <- 1 - pf(chow, m, n-2*m)
  out <- list(SSE.1=SSE.1, SSE.2=SSE.2, SSE.rm=SSE.rm, chow=chow, pval.chow=pval.chow)
  return(out)
})
sapply(chow.test, cbind)
```

```{r eval=FALSE, include=FALSE}
#log working in progress. Courrently not working
attach(dat1)
na.action="na.omit"
#dat1[dat1 == "0"] <- 0.01
TDC2[which(is.nan(TDC2))] = 0.01
TDC2[which(is.na(TDC2))] = 0.01
TDC2[which(TDC2==Inf)] = 0.01
# ANOVA
baseMod <- lm(log(TDC2) ~ AGE + log(BONUS) + log(AT) + log(CH) + log(DLTT) + log(NI) + log(PRCH_C) + log(PRCL_C))
mod1 <- lm(log(TDC2) ~ AGE + log(BONUS) + log(AT) + log(CH) + log(DLTT) + log(NI))
mod2 <- lm(log(TDC2) ~ AGE + log(BONUS) + log(AT) + log(CH) + log(NI))
mod3 <- lm(log(TDC2) ~ AGE + log(BONUS) + log(AT) + log(CH) + log(DLTT))
mod4 <- lm(log(TDC2) ~ AGE + log(BONUS) + log(AT) + log(CH))
mod5 <- lm(log(TDC2) ~ AGE + log(BONUS) + log(AT))
anova(baseMod, mod1, mod2, mod3, mod4, mod5)

##na.action=na.exclude,data=dat
```


